---
title: "Ciccolini_Lab4"
author: "Skylar Ciccolini"
date: "2025-09-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Set Up

```{r}
library(pacman)

p_load(tidyverse, sf, spData)

```


## 2. Data

```{r}

data("world")
data("worldbank_df")

```


## 3. Assignment

### 3.1 Examine the data

```{r}
glimpse(world)

class(world)

glimpse(worldbank_df)

class(worldbank_df)

```

> world has multipolygon data and two different subclasses of dataframes. worldbank doesn't have any spatial data.

Combine dataframes

```{r}

worldwide <- left_join(world, worldbank_df, by = c("iso_a2"))

glimpse(worldwide)

```

> The resulting dataframe looks reasonable to me. All the data I expected to be there is.


### 3.2 Make some basic (non-spatial) plots to explore differences in urban population and literacy across countries.

```{r}

ggplot() +
  geom_point(data = worldwide, aes(x = iso_a2, y = urban_pop))

ggplot() +
  geom_point(data = worldwide, aes(x = iso_a2, y = literacy))

```


### 3.3 Now make some basic maps to explore urban population and literacy across countries.

```{r}

ggplot() +
  geom_sf(data = worldwide, aes(fill = urban_pop))

ggplot() +
  geom_sf(data = worldwide, aes(fill = literacy)) ### this seems hinky...

```



### 3.4 Make some basic (non-spatial) plots to explore differences in urban population and literacy across continents.

```{r}

ggplot() +
  geom_point(data = worldwide, aes(x = continent, y = urban_pop))

ggplot() +
  geom_point(data = worldwide, aes(x = continent, y = literacy))

```



### 3.5 Now make some basic maps to explore differences in urban population and literacy across continents.
Change to a different, projected coordinate system for these maps.

```{r}

# change the PCS to Equal earth

worldwide_proj <- st_transform(worldwide, "+proj=eqearth +lon_0=0 +datum=WGS84 +units=m +no_defs
")

# now, to plot
ggplot() +
  geom_sf(data = worldwide_proj, aes(fill = urban_pop))

ggplot() +
  geom_sf(data = worldwide_proj, aes(fill = literacy)) ### this seems hinky...
```




























